---
title: "Portland"
author: "Sandhya Kajeepeta"
date: "2024-05-07"
output: html_document
---
```{r load packages,echo=FALSE, results=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggplot2)
library(sf)
library(readxl)
library(pscl)
library(nlme)
```

#Portland

```{r}
c16 <- read_csv("./Data/Portland/CrimeData-2016.csv")
c17 <- read_csv("./Data/Portland/CrimeData-2017.csv")
c18 <- read_csv("./Data/Portland/CrimeData-2018.csv")
```

```{r}
df <- rbind(c16, c17, c18)
df$OccurDate <- as.POSIXct(df$OccurDate, format="%m/%d/%Y")
df$OccurMonth <- format(df$OccurDate, "%Y-%m")
df <- df %>% filter(OccurDate > "2016-10-14" & OccurDate < "2018-10-16")
```

```{r}
pmonthly <- df %>% group_by(OccurMonth) %>% summarize(total_crime = sum(OffenseCount), person_crime=sum(OffenseCount[CrimeAgainst=="Person"]), prop_crime=sum(OffenseCount[CrimeAgainst=="Property"]))

pdaily <- df %>% group_by(OccurDate) %>% summarize(total_crime = sum(OffenseCount), person_crime=sum(OffenseCount[CrimeAgainst=="Person"]), prop_crime=sum(OffenseCount[CrimeAgainst=="Property"]))

```

```{r}
#plot monthly crime

ggplot(data=pmonthly, mapping=aes(x=OccurMonth, y=total_crime, group=1)) +
  geom_line() +
  theme_bw() + theme(legend.key=element_blank()) +
  ylab("Total reported crime offenses")
```

```{r}
#code date of gang database abolition
dateAbolish <- as.Date("2017-10-15")

pdaily$postAbolish <- (pdaily$OccurDate >= dateAbolish)

pdaily$DateNum <- as.numeric(as.Date(pdaily$OccurDate))

pdaily$DateNumCtr <- pdaily$DateNum - as.numeric(dateAbolish)
```

##Total crime
Linear model fit for correlogram
```{r}
lmFit1 <- lm(total_crime ~ DateNum, data=pdaily)
acf(resid(lmFit1))
```

Crude ITS
```{r}

glsFit1 <- gls(model = total_crime ~ DateNumCtr + postAbolish + DateNumCtr:postAbolish,
               data = pdaily,
               correlation = corARMA(p=7))
summary(glsFit1)
```
Overplot prediction
```{r}
#Create prediction dataset
newdata <- data.frame(DateNumCtr = seq(min(pdaily$DateNumCtr), max(pdaily$DateNumCtr), by=1))
newdata$postAbolish <- (newdata$DateNumCtr >=0)

#Predict
newdata$total_crime <- predict(glsFit1, newdata=newdata)

#Join dates back
temp <- pdaily %>% dplyr::select(DateNumCtr, OccurDate)
newdata <- left_join(newdata, temp, by="DateNumCtr")

ggplot(data=pdaily, mapping=aes(x=as.Date(OccurDate), y=total_crime, group=1)) +
  geom_line() +
  geom_line(data=subset(newdata, DateNumCtr < 0), color="red", size=1.5) +
  geom_line(data=subset(newdata, DateNumCtr >=0), color="red", size=1.5) +
  geom_vline(xintercept=as.numeric(as.Date("2017-10-15")), linetype=2) +
  theme_bw() + theme(legend.key=element_blank()) +
  ylab("Daily total crime offenses") +
  xlab(" ") +
  scale_x_date(date_breaks="4 months", date_labels = "%b %Y")
ggsave("Portland_total.png", width = 5, height = 3, units = "in")
```
##Violent crime
Linear model fit for correlogram
```{r}
lmFit1 <- lm(person_crime ~ DateNum, data=pdaily)
acf(resid(lmFit1))
```

Crude ITS
```{r}

glsFit1 <- gls(model = person_crime ~ DateNumCtr + postAbolish + DateNumCtr:postAbolish,
               data = pdaily,
               correlation = corARMA(p=7))
summary(glsFit1)
```
Overplot prediction
```{r}
#Create prediction dataset
newdata <- data.frame(DateNumCtr = seq(min(pdaily$DateNumCtr), max(pdaily$DateNumCtr), by=1))
newdata$postAbolish <- (newdata$DateNumCtr >=0)

#Predict
newdata$person_crime <- predict(glsFit1, newdata=newdata)

#Join dates back
temp <- pdaily %>% dplyr::select(DateNumCtr, OccurDate)
newdata <- left_join(newdata, temp, by="DateNumCtr")

ggplot(data=pdaily, mapping=aes(x=as.Date(OccurDate), y=person_crime, group=1)) +
  geom_line() +
  geom_line(data=subset(newdata, DateNumCtr < 0), color="red", size=1.5) +
  geom_line(data=subset(newdata, DateNumCtr >=0), color="red", size=1.5) +
  geom_vline(xintercept=as.numeric(as.Date("2017-10-15")), linetype=2) +
  theme_bw() + theme(legend.key=element_blank()) +
  ylab("Daily violent crime offenses") +
  xlab(" ") +
  scale_x_date(date_breaks="4 months", date_labels = "%b %Y")
ggsave("Portland_violent.png", width = 5, height = 3, units = "in")
```

##Property crime
Linear model fit for correlogram
```{r}
lmFit1 <- lm(prop_crime ~ DateNum, data=pdaily)
acf(resid(lmFit1))
```

Crude ITS
```{r}

glsFit1 <- gls(model = prop_crime ~ DateNumCtr + postAbolish + DateNumCtr:postAbolish,
               data = pdaily,
               correlation = corARMA(p=7))
summary(glsFit1)
```
Overplot prediction
```{r}
#Create prediction dataset
newdata <- data.frame(DateNumCtr = seq(min(pdaily$DateNumCtr), max(pdaily$DateNumCtr), by=1))
newdata$postAbolish <- (newdata$DateNumCtr >=0)

#Predict
newdata$prop_crime <- predict(glsFit1, newdata=newdata)

#Join dates back
temp <- pdaily %>% dplyr::select(DateNumCtr, OccurDate)
newdata <- left_join(newdata, temp, by="DateNumCtr")

ggplot(data=pdaily, mapping=aes(x=as.Date(OccurDate), y=prop_crime, group=1)) +
  geom_line() +
  geom_line(data=subset(newdata, DateNumCtr < 0), color="red", size=1.5) +
  geom_line(data=subset(newdata, DateNumCtr >=0), color="red", size=1.5) +
  geom_vline(xintercept=as.numeric(as.Date("2017-10-15")), linetype=2) +
  theme_bw() + theme(legend.key=element_blank()) +
  ylab("Daily property crime offenses") +
  xlab(" ") +
  scale_x_date(date_breaks="4 months", date_labels = "%b %Y")
ggsave("Portland_property.png", width = 5, height = 3, units = "in")
```


#Chicago

```{r}
c <- read_csv("./Data/Chicago/Crimes_-_2001_to_Present.csv")
c$Date <- as.POSIXct(c$Date, format="%m/%d/%Y %H:%M:%S")
c$OccurDate <- as.Date(format(c$Date, "%Y-%m-%d"))
```

```{r}
#define person/violent and property
c$CrimeAgainst <- ifelse(c$`Primary Type` %in% c("ASSAULT","BATTERY","CRIMINAL SEXUAL ASSAULT","HOMICIDE","HUMAN TRAFFICKING","KIDNAPPING","OFFENSE INVOLVING CHILDREN","SEX OFFENSE","STALKING"), "Person", ifelse(c$`Primary Type` %in% c("ARSON","BURGLARY","CRIMINAL DAMAGE","DECEPTIVE PRACTICE","MOTOR VEHICLE THEFT","ROBBERY","THEFT"), "Property", NA))
```


```{r}
cdaily <- c %>% group_by(OccurDate) %>% summarise(total_crime = n(), person_crime=sum(CrimeAgainst=="Person", na.rm=T), prop_crime=sum(CrimeAgainst=="Property", na.rm=T))

```

```{r}
#code date of gang database abolition
dateAbolish <- as.Date("2023-09-07")

cdaily$postAbolish <- (cdaily$OccurDate >= dateAbolish)

cdaily$DateNum <- as.numeric(cdaily$OccurDate)

cdaily$DateNumCtr <- cdaily$DateNum - as.numeric(dateAbolish)

cdaily <- cdaily %>% filter(!is.na(OccurDate) & OccurDate!="2024-04-29")
```

##Total crime
Linear model fit for correlogram
```{r}
lmFit1 <- lm(total_crime ~ DateNum, data=cdaily)
acf(resid(lmFit1))
```

Crude ITS
```{r}

glsFit1 <- gls(model = total_crime ~ DateNumCtr + postAbolish + DateNumCtr:postAbolish,
               data = cdaily,
               correlation = corARMA(p=1))
summary(glsFit1)
```
Overplot prediction
```{r}
#Create prediction dataset
newdata <- data.frame(DateNumCtr = seq(min(cdaily$DateNumCtr), max(cdaily$DateNumCtr), by=1))
newdata$postAbolish <- (newdata$DateNumCtr >=0)

#Predict
newdata$total_crime <- predict(glsFit1, newdata=newdata)

#Join dates back
temp <- cdaily %>% dplyr::select(DateNumCtr, OccurDate)
newdata <- left_join(newdata, temp, by="DateNumCtr")

ggplot(data=cdaily, mapping=aes(x=as.Date(OccurDate), y=total_crime, group=1)) +
  geom_line() +
  geom_line(data=subset(newdata, DateNumCtr < 0), color="red", size=1.5) +
  geom_line(data=subset(newdata, DateNumCtr >=0), color="red", size=1.5) +
  geom_vline(xintercept=as.numeric(as.Date("2023-09-07")), linetype=2) +
  theme_bw() + theme(legend.key=element_blank()) +
  ylab("Daily total crime offenses") +
  xlab(" ") +
  scale_x_date(date_breaks="4 months", date_labels = "%b %Y")
ggsave("Chicago_total.png", width = 5, height = 3, units = "in")
```
##Violent crime
Linear model fit for correlogram
```{r}
lmFit1 <- lm(person_crime ~ DateNum, data=cdaily)
acf(resid(lmFit1))
```

Crude ITS
```{r}

glsFit1 <- gls(model = person_crime ~ DateNumCtr + postAbolish + DateNumCtr:postAbolish,
               data = cdaily,
               correlation = corARMA(p=1))
summary(glsFit1)
```
Overplot prediction
```{r}
#Create prediction dataset
newdata <- data.frame(DateNumCtr = seq(min(cdaily$DateNumCtr), max(cdaily$DateNumCtr), by=1))
newdata$postAbolish <- (newdata$DateNumCtr >=0)

#Predict
newdata$person_crime <- predict(glsFit1, newdata=newdata)

#Join dates back
temp <- cdaily %>% dplyr::select(DateNumCtr, OccurDate)
newdata <- left_join(newdata, temp, by="DateNumCtr")

ggplot(data=cdaily, mapping=aes(x=as.Date(OccurDate), y=person_crime, group=1)) +
  geom_line() +
  geom_line(data=subset(newdata, DateNumCtr < 0), color="red", size=1.5) +
  geom_line(data=subset(newdata, DateNumCtr >=0), color="red", size=1.5) +
  geom_vline(xintercept=as.numeric(as.Date("2023-09-07")), linetype=2) +
  theme_bw() + theme(legend.key=element_blank()) +
  ylab("Daily violent crime offenses") +
  xlab(" ") +
  scale_x_date(date_breaks="4 months", date_labels = "%b %Y")
ggsave("Chicago_violent.png", width = 5, height = 3, units = "in")
```

##Property crime
Linear model fit for correlogram
```{r}
lmFit1 <- lm(prop_crime ~ DateNum, data=cdaily)
acf(resid(lmFit1))
```

Crude ITS
```{r}

glsFit1 <- gls(model = prop_crime ~ DateNumCtr + postAbolish + DateNumCtr:postAbolish,
               data = cdaily,
               correlation = corARMA(p=1))
summary(glsFit1)
```
Overplot prediction
```{r}
#Create prediction dataset
newdata <- data.frame(DateNumCtr = seq(min(cdaily$DateNumCtr), max(cdaily$DateNumCtr), by=1))
newdata$postAbolish <- (newdata$DateNumCtr >=0)

#Predict
newdata$prop_crime <- predict(glsFit1, newdata=newdata)

#Join dates back
temp <- cdaily %>% dplyr::select(DateNumCtr, OccurDate)
newdata <- left_join(newdata, temp, by="DateNumCtr")

ggplot(data=cdaily, mapping=aes(x=as.Date(OccurDate), y=prop_crime, group=1)) +
  geom_line() +
  geom_line(data=subset(newdata, DateNumCtr < 0), color="red", size=1.5) +
  geom_line(data=subset(newdata, DateNumCtr >=0), color="red", size=1.5) +
  geom_vline(xintercept=as.numeric(as.Date("2023-09-07")), linetype=2) +
  theme_bw() + theme(legend.key=element_blank()) +
  ylab("Daily property crime offenses") +
  xlab(" ") +
  scale_x_date(date_breaks="4 months", date_labels = "%b %Y")
ggsave("Chicago_property.png", width = 5, height = 3, units = "in")
```

